Impact of grids and dynamical cores in CESM2.2 on the surface mass balance of the Greenland Ice Sheet

### 2022-09-06 first_review/response-to-reviewers.tex

\documentclass[12pt,oneside,a4paper]{article}%{amsart}
\usepackage[margin=1in]{geometry} % see geometry.pdf on how to lay out the page. There's lots.
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{enumitem}
\usepackage{caption}
\setlength\parindent{0pt}
\usepackage[round]{natbib}
\bibliographystyle{abbrvnat}
\renewcommand{\bibsection}{}

\title{\large Response to Reviewers}
\author{\normalsize Adam R. Herrington and CoAuthors}
\date{} % delete this line to display the current date

\begin{document}

\maketitle

\section*{General Remarks}

{\color{red}{We would like to thank both reviewers for taking the time to provide thoughtful feedback on our manuscript.}} \newline

\section*{Reviewer 1}

\subsection*{General comments}

Given that the different grid and dycore choices come with considerable different computational costs, I think it would be useful to include a section addressing that specific question. Some information touching on the costs appears at different places in the manuscript. It would be good to bring that all together somewhere and make a clear comparison, e.g. including a comparative table. \newline

The way the term surface mass balance (SMB) is defined and used in the manuscript is confusing. The different components to the SMB are not consistently applied and there are two different definitions of the SMB in use. In part this confusion is related to the fact that the model itself does not calculate all the SMB components and assumptions have to be made about the separation between snow and ice, especially in the currently discussed uncoupled case. I would suggest to start with a physical definition of SMB with all its components in an equation and then explain based on that what the model is actually doing and what constraints have to be applied. In this context it may help to describe what the model would do in a coupled setup with active CISM and then apply further constraints for the diagnostic case currently described. \newline

In the discussion about the different dycores, I was wondering if there could be a notion that the latitude-longitude approaches are more mature because they have a longer history. Maybe further improvements on technical aspects like filtering and optimal solutions may be expected for the FE approaches. \newline

The relationship between atmospheric grid and land grid seems to be important. Would it be possible to say that higher resolution of the atmospheric grid improves the precipitation, while increased resolution of the land model improves the melt calculation/ablation components? \newline

\subsection*{Specific comments}

l12 I am confused about the use of the terms "uniform resolution" and "equivalent resolution". Since the lat-lon grids are also uniform (they have 1 or 2 degree resolution everywhere) I find this is not a good description to distinguish them here. Maybe this requires a few more words to make it clear to the reader what is meant here. The confusion is not fully removed in the main text either. See later comments.\newline

{\color{blue}{We take your point and agree that this is confusing. To clarify this, we've replaced ``uniform resolution" with ``quasi-uniform unstructured grids" and clarify that ``uniform" refers to isotropic grids where every grid cell has the same length in meters, as opposed to an angular distance in spherical coordinates. Whereas we had previously used ``uniform grids" to refer to both lat-lon and quasi-uniform unstructured grids, we replace ``uniform" with ``lat-lon and quasi-uniform grids" throughout the manuscript, to be more explicit about the differences between the SE and lat-lon grids.}} \newline

l29 It would be good to add one sentence about the cost here. E.g. "... recovering resolution lost in the transition to quasi-uniform grids, albeit at increased computational cost." \newline

{\color{blue}{Done.}} \newline

l37 "latitude-longitude grids" \newline

{\color{blue}{Done.}} \newline

l75 "Dycores built on lat-lon grids have some advantages over dycores built on unstructured grids" \newline

{\color{blue}{Done.}} \newline

l92 "have large" --> "has a large" \newline

{\color{blue}{Done.}} \newline

l94 The logic of "Conversely" is not clear, since I think it is referring to the first sentence in the paragraph not the preceding one. Reformulate? \newline

{\color{blue}{Sentence beginning with conversely has been reformulated to improve clarity.}} \newline

l107 Move "(SMB; the integrated sum of precipitation and runoff)" before "of the Greenland ice sheet". \newline

{\color{blue}{Done.}} \newline

l104-110 With the benefit of a high resolution land model in mind, is it possible to run the land model at a different/higher resolution than the atmosphere model? If so, this could be mentioned here. \newline

{\color{blue}{As discussed above, I'm not optimistic about the utility of a higher-res land model, and would rather not go on a tangent about custom configurations.}} \newline

l117 "(marginal regions where seasonal melting exceeds the annual mass input from precipitation)" This sentence is not correct. What defines the ablation area is a negative SMB. Problems with the statement are that melt can refreeze and precipitation can occur as snow and rain. It would maybe help to write out the whole SMB in an equation if there is confusion about these terms. \newline

{\color{blue}{Replaced with ``marginal regions where the annual SMB is negative."}} \newline

l138 "is represented with 1/4 resolution" I believe they have used two different levels of refinement (0.5 and 0.25 degree): Maybe state "is represented with up to 1/4 resolution" instead. Or give resolution in km as they do in most of the paper: 55 km and 28 km. \newline

{\color{blue}{Used reviewers ``up to" suggestion.}} \newline

l159 What I miss in this section is a short introduction to CESM. Can be brief, but should bring the reader up to speed what kind of model that is \newline

{\color{blue}{Included a short introduction to CESM. Also moved language from the experimental design section in here, as well as include a statement that the ice sheet model is not dynamically active. Our hope is that by including this at the very top of the methods section, this partly addresses the reviewers concern that it is not clear we are not running with an active ice sheet.}} \newline

l174 Is it correct to say the dycore is a model? \newline

{\color{blue}{Reworded.}} \newline

l203 "the SE numerics have been enhanced". Enhanced compared to what. An earlier model version, I suspect. Mention that. \newline

{\color{blue}{added: relative to the CESM2.0 release.}} \newline

l214 "... has been modified to scale the smoothing radius by the local element size, resulting in rougher topography in the refinement zone." This seems in contrast to the very idea of refinement. Maybe this should be mentioned here? \newline

{\color{blue}{I don't follow. How is rougher topography in the refinement zone inconsistent with the very idea of refinement? From our perspective it is a requirement of refinement -- finer dynamics permits rougher topography.}} \newline

l215 "For SE grids". The grids are not SE! It is clear what you mean, but good to use proper terminology. Maybe "When using the SE dycore with quasi-uniform element sizes ..." \newline

{\color{blue}{Used the reviewers preferred wording.}} \newline

l215-219 I seem to understand from this paragraph that CSLAM is an improvement, but that it cannot be applied for VR grids. If so, please make that clear here. \newline

{\color{blue}{Added. Yes, this is unfortunately true. We do have plans to address this in VR within the next 2 years.}} \newline

l220-224 I am wondering if this separate grid could in principle be extended to match an underlying SE dynamics VR grid by further subdivision into more control volumes. \newline

{\color{blue}{The underlying issue is that the SE dynamics don't have control volumes. They have GLL nodes, which are point samples of a spectrally truncated state. In order to couple SE to other model components, we had to invent control volumes because the CESM infrastructure uses finite-volume mapping algorithms to couple to other model components. The resulting control volume grid is an anisotropic mess (see Figure 4 in my MWR 2019 paper), which was constructed by constructing control volumes consistent with each GLL node's quadrature weight. It is much preferable to split the elements into quadrilateral control volumes and use the continuous basis representation to project the SE solution onto the quadrilateral control volumes (like we do in pg3 and pg2.}} \newline

l224 add "at" after "evaluated" \newline

{\color{blue}{Done.}} \newline

l249 I don't think "RACMO" has been defined yet. Please do. \newline

{\color{blue}{Done.}} \newline

l251 The discussion about computational costs is an important one in this comparison and should get at least an own paragraph, if not an own sub-section. \newline

{\color{blue}{See introductory remarks.}} \newline

l288 It is not the grids that calculate the SMB. Maybe "CESM simulates the GrIS SMB as the sum of mass accumulation and ablation" \newline

{\color{blue}{We have replaced with the suggested language. In order to reconcile your following comments regarding this paragraph, what follows this top sentence is a description of how the SMB is computed in CESM. This requires some changes to the language, such as excluding rain from the runoff terms, and making clear only ice accumulation is contained in the accumulation terms. Having addressed some unclear statements about the definition of the SMB in the introduction, we don't feel that we need to add another equation for CESM's SMB, or a common glaciological definition plus a simplified version used by CESM. These equations can be found in the references in the sentence "For a more detailed description of the SMB is computed in CESM ...", in particular equations 4 and 5 in van Kampenhout et al. 2020. }} \newline

l289 Strictly speaking only snow accumulation is a definite positive mass source, while rain can either runoff or refreeze. Again, an equation that gives all the terms in the SMB may be useful here. \newline

{\color{blue}{See above comment.}} \newline

l290 Suggest to replace "with runoff being a combination of" by "with runoff having sources of a combination of". Again, I think runoff is not a straight sum of these, because of refreezing. \newline

{\color{blue}{This has been modified in the reworked section.}} \newline

l300 Mention here that this can change if the Greenland ice sheet is fully coupled. I generally miss already earlier a statement that you are using CESM without interactive Greenland ice sheet. \newline

{\color{blue}{This is a good catch and the language has been inserted about how this changes with an active ice sheet.}} \newline

l305 These descriptions apply to the uncoupled case, which should be clarified. It may also be interesting to explain what happens in the coupled case, since CESM has that as an option. I am a bit surprised that this is not the primary option you are describing here. \newline

{\color{blue}{After addressing your last comment, we believe the paragraph is valid for both coupled and un-coupled cases.}} \newline

l308 I think it would be interesting to show that downscaled SMB (using the elevation classes) and compare that between the different cases. That is the SMB product one would actually use either in the coupled case or when extracting the SMB from CESM for offline forcing of an ice sheet model. In addition, you could compare that result more easily between the different cases as they would all be well-defined on the output grid. \newline

{\color{blue}{We agree that would be interesting. The reason we didn't employ this approach -- which to be clear, we interpret as evaluating the SMB field downscaled to the 4km CISM grid -- is that only the SMB is downscaled. It is the resolution sensitivity of the individual components of the SMB that is the main focus of this study. It's representation on the CLM grid is still quite advanced, owing to the EC scheme. I suppose we could have tried to hack the code to downscale the SMB components to CISM, but we didn't go down that road (and no telling the number of issues that would've cropped up going down that road). That said, our approach of mapping to the common ice mask is quite complicated as well.}} \newline

l313 "Changes in ice depth, but not snow depth, count toward the SMB" seems to be in contrast with "Any snow above the 10 m cap contributes towards ice accumulation in the SMB" in the paragraph before. Please clarify. \newline

{\color{blue}{I'm not sure I see the contradiction the reviewer is referring to, but this a clearly a redundancy that we've eliminated by consolidating the paragraphs.}} \newline

l322 The datasets you use are not observational but model results. I think this should be made clear already in the heading. Maybe "Target" or "Validation" datasets. \newline

{\color{blue}{Done.}} \newline

l323 Again, the datasets in use are not observational. \newline

{\color{blue}{Done.}} \newline

l323-338 You do not seem to describe anywhere the other datasets in Table 2: ERA, CERES, Calipso. These should all be described here as well. \newline

{\color{blue}{Done.}} \newline

l331 add "e.g." after before "maintained". This is a very specific collection of the data. \newline

{\color{blue}{I'm not sure I follow, but I did clarify the second category to be less specific (remote sensing datasets instead of radar accumulation datasets).}} \newline

l343 "reference dataset" Remind us what this is and where it is sourced from. \newline

{\color{blue}{Clarified the reference dataset and added citation.}} \newline

l348 "can lead to large differences in integrated SMB" Could explain briefly why that is, because the mass loss happens in a narrow band along the margins. \newline

{\color{blue}{The Hansen paper indicates that differences in the ice masks occur in its representation of the ice margins, where lots of melting conceivable occurs. So an erroneously small ice mask may overestimate the SMB due to omitting areas where melting occurs. }} \newline

l363 Equ 1. Could be read as "evaporation divided by sublimation". Why not "SMB = accumulation + runoff + evaporation + sublimation"? \newline

{\color{blue}{Done.}} \newline

l369 This is confusing, because we are in section 2.4. If you are operating with two different SMB terms, you should have two equations to clearly distinguish them. \newline

{\color{blue}{We have reworded and moved this last sentence, indicating that the total water SMB is different than the SMB internally computed by CLM, as a topic sentence to its own paragraph immediately following. We've added clarification on the role of the internally generated CLM SMB, that we don't use this definition in this study.}} \newline

l372 "grid-cell mean quantities" Clarify that these are atmospheric/land-model grid cells. \newline

{\color{blue}{Apologies for not understanding the suggestion here. What is it that needs clarification? A topic sentence in a paragraph in the methods section begins with ``CLM runs on the same grid as the atmosphere..." Is that what you are asking about?}} \newline

l388 insert "further" before "diversify". \newline

{\color{blue}{Done.}} \newline

l439 Since you start the section in 404 with "Before delving into the simulated Arctic climate conditions ...", could mark that we are now moving from global to Arctic discussion. \newline

{\color{blue}{Good suggestion. Done}} \newline

l461 "CERES product" needs to be described in 2.5. \newline

{\color{blue}{Done.}} \newline

l478 Add "radiation" after "shortwave". \newline

{\color{blue}{Done.}} \newline

l525 "underline"? \newline

{\color{blue}{Yes. Fixed.}} \newline

l528 Did you test how much biased the results could be due to the relatively short period of 20 years in view of the large inter-annual variability? If you don't have more years, you could remove any of the 20 years and see how much the numbers change. \newline

l553 "smaller melt-rates" or smaller melt-rate biases? \newline

{\color{blue}{since both are true, replaced with ``smaller melt-rates and melt-rate biases."}} \newline

l564 Often PDD schemes have different PDD factors for snow and ice melt. Is this not not the case here? \newline

{\color{blue}{To simplify the calculation I just took an intermediate value between snow and ice (Braithwaite 1984 uses 3 mm/d/C and 7 mm/d/C for ice), so I wouldn't need to approximate how to partition at particular monthly average grid cell. Added these details in the manuscript in parenthesis.}} \newline

l557-564 Description and motivation for using the PDD model should probably be moved/added to the Methods section. \newline

{\color{blue}{Our personal opinion is that defining it on the fly, naturally arising from the analysis to explain a discrepancy between two contrasting results, is preferable instead of premeditating it earlier on in the manuscript.}} \newline

l571 Consider reordering figures, since we are jumping back from 11 to 9 here. \newline

l594 This reads like a contradiction to me "coarser models ... depress melt rates, which is the opposite of the melt bias that occurs in the coarse grid simulations". What happens in course models is the opposite to what happens in coarse models? \newline

{\color{blue}{Reworded this slightly to improve the clarity. ``This suggests that coarser models will tend to elevate the ablation zones relative to where they should be, which may be expected to cause anomalous (adiabatic) cooling and depressed melt rates, but this is opposite the melt bias that occurs in the coarse grid simulations."}} \newline

l603 "volume averaging" I would have guessed this was done by area averaging. \newline

{\color{blue}{Fair point. Changed to area averaging.}} \newline

l605 Maybe add "and more melting" after "warm bias over the ice-covered patch". \newline

{\color{blue}{Done.}} \newline

l622 Define PDFs. \newline

{\color{blue}{Done.}} \newline

l648 Add "refined over the US" after CONUS. \newline

{\color{blue}{Done.}} \newline

l735 hydrid --> hybrid \newline

{\color{blue}{Done.}} \newline

l748 Replace "shown" by "given" to avoid repeat of "showing/show". \newline

{\color{blue}{Done.}}

\subsection*{Figures}

\subsubsection*{Fig 1}
I think the grid lines could be thinner. On a printout this doesn't come out well. \newline
The color choice puts a lot of emphasis on the lakes in blue. Suggest to display the land-sea mask instead, i.e. make the lakes part of the land mask. \newline

{\color{blue}{Thinned the lines by half and changed the lake colors to tan/land.}} \newline

I would also consider using another projection that includes all of Greenland and it's surroundings. \newline

{\color{blue}{We have decided to keep the original polar projection, in order to show the polar points and Greenland.}} \newline

Caption. It is confusing to refer to all 4 grids as uniform, because they are uniform in different ways. At least distinguish them as uniform 1, 2 latitude-longitude grids (top) and quasi-uniform element size grids (bottom), or similar. \newline

{\color{blue}{Changed to new labeling discussed above.}}

\subsubsection*{Fig 10}
Better use y-axis label ICE+SNOW MELT consistent with Fig 11 and to avoid confusion ICE divided by SNOW. \newline

\subsubsection*{Fig 11}
The sign convention seems to be changed from Fig 10 with positive melt. Make consistent. \newline

\subsection*{References}

Many references are not formatted correctly in terms of upper case / lower case. There are also other problems like incomplete author lists and formatting problems. Please revise. Some examples below. \newline

l793 This is a regular paper citation that doesn't need a "Retrieved from" entry. Author list incomplete. \newline

l797 Incomplete author list. \newline

l924 Refer to final paper. Note title change: NCAR Topo v1.0 NCAR global model topography generation software for unstructured grids. https://doi.org/10.5194/gmd-8-3975-2015 \newline

l988 This is a regular paper citation that doesn't need a "Retrieved from" entry. Author list with "...". \newline

l1063 Refer to final paper. https://doi.org/10.5194/gmd-14-5023-2021 \newline

l1071 Refer to final paper. https://doi.org/10.5194/tc-13-1547-2019 \newline

\section*{Reviewer 2}

%\section*{\normalsize References}
%\setlength{\bibsep}{0pt}
%{\footnotesize
%\bibliography{bib}}

\end{document}

### 2022-09-06 first_submit/arcticgrids.tex
first commit after review

\begin{keypoints}
\item The CESM2.2 release includes several enhancements to the spectral-element dynamical core, including two new Arctic refined mesh configurations.
\item Uniform resolution grids degrade the surface mass balance of the Greenland Ice Sheet compared with equivalent low resolution latitude-longitude grids.
\item The refined Arctic meshes substantially improve the surface mass balance over all low resolution grids.
\end{keypoints}

@@

%\section*{Plain Language Summary}
%[ enter your Plain Language Summary here or delete this section]

@@

%{\color{blue} (WHL: Would it be worth mentioning that quasi-uniform grids also have benefits in polar regions, e.g. by removing elongated grid cells?)}{\color{red} (ARH: I don't think it's worth getting into that here. You'll see later in the manuscript that b/c the cells are narrower in the E-W direction, more cells can fit into the big ablation zone on the western margin.)}{\color{purple}(PHL: another advantage of lat-lon grids is that the coordinate lines are orthogonal!)}

%{\color{blue} (WHL: This raises the question: How do lat-lon grids better resolve polar regions at a given nominal global resolution, if the filtering makes them effectively coarser?  Does the damping degrade some processes more than others?)}{\color{red} (ARH - The phrase "approximately the same" doing a lot of heavy lifting here b/c I frankly don't know. The results from this paper suggest that the effective resolution is finer over polar regions, but this is a bit hand-wavy, despite my attempts to quantify it more exactly, and based on the precipitation rates having smaller biases. Also, to be clear, effective resolution is specific to atmospheric features; the land model is free to take advantage of these additional grids cells representing the ablation zone, and they certainly do. Perhaps I can add something to the effect of ``approximately the same everywhere on the grid, although marginal increases in effective resolution over the polar regions cannot be ruled out...")\\

%{\color{purple}PHL: These are good points! I think it is useful to separate dynamics from physics in this discussion. Physics is called at more columns on the lat-lon grid as one approaches the poles. There is no "polar filtering" in physics so effectively the physics resolution is much higher than on a the cubed-sphere grid as one approaches the mid-latitudes and poles. Dynamics, on the other hand, filters out smaller scales so effectively it should only retain the 1 degree scale. That said, you will get a more accurate solution since you have more grid-points to represent waves (even though smaller scales are filtered) compared to the cubed-sphere grid. I think that is pretty clear in flow over orography. One can compare the lat-lon grid setup to a spectral model where one keeps viscosity coefficients fixed as one increased horizontal resolution - in such a setup one will get convergence of the dynamics!}}

\begin{figure}[t]
\begin{center}
\begin{tabular}{cc}
         \includegraphics[width=50mm]{figs/grid-f19.pdf}&
         \includegraphics[width=50mm]{figs/grid-f09.pdf}\\
         \includegraphics[width=50mm]{figs/grid-ne30pg2.pdf}&
         \includegraphics[width=50mm]{figs/grid-ne30pg3.pdf} \\
\end{tabular}
\end{center}
\caption{Computational grids for the uniform $1^{\circ}-2^{\circ}$ grids in this study. Grids names after Table~\ref{tbl:table1}, (a) $\tt f19$, (b) $\tt f09$, (c) $\tt ne30pg2$ and (d) $\tt ne30pg3$.}
\label{fig:uni-grids}
\end{figure}

\begin{figure}[t]
\begin{center}
\begin{tabular}{ccc}
         \includegraphics[width=40mm]{figs/grid-ARCTIC.pdf}&
         \includegraphics[width=40mm]{figs/grid-ARCTICGRIS.pdf}&
         \includegraphics[width=40mm]{figs/grid-CONUS.pdf} \\
\end{tabular}
\end{center}
\caption{Variable-resolution grids available in CESM2.2, (a) $\tt Arctic$, (b) $\tt Arctic-GrIS$ and (c) $\tt CONUS$. Note what is shown is the element grid; the computational grid has $3\times3$ independent grid points per element.}
\label{fig:vr-grids}
\end{figure}

@@

\begin{figure}[t]
\begin{center}
         \includegraphics[width=100mm]{figs/temp_grisres.pdf}
\end{center}
\caption{The spatial properties of the GrIS as represented by different grids in this study. (Left) GrIS area error, computed as the relative differences from a 4-km dataset used to create the CESM ice masks, (right) approximate average grid spacing over GrIS.}
\label{fig:grisdx}
\end{figure}

@@

\begin{figure}[t]
\begin{center}
         \includegraphics[width=130mm]{figs/temp_dhgt_panel_STEND_CLUBB-lores.pdf}
\end{center}
\caption{1979-1998 annual mean temperature (left column) and CLUBB temperature tendencies (right column) in zonal mean height space, expressed as differences between the various $1^{\circ}-2^{\circ}$ grids. The thick grey and magenta lines are the tropopause for the control run and the test run, respectively.}
\label{fig:dT-lores}
\end{figure}

@@

\begin{figure}[t]
\begin{center}
         \includegraphics[width=130mm]{figs/temp_dhgt_panel_STEND_CLUBB-hires.pdf}
\end{center}
\caption{As in Figure~\ref{fig:dT-lores} but for the short-time-step experiment and the VR grids. The fields plotted are for the climatological northern hemisphere summer. We focus on summer because that is when the resolution response is largest, and the refined regions are located in the northern hemisphere.}
\label{fig:dT-hires}
\end{figure}

@@

\begin{figure}[t]
\begin{center}
         \includegraphics[width=100mm]{figs/temp_contours_diffERA5_Thyps.pdf}
\end{center}
\caption{1979-1998 lower troposphere, northern hemisphere summer virtual temperature biases, computed as the difference from ERA5. Lower troposphere layer mean virtual temperature is derived from the 1000 hPa - 500h Pa geopotential thickness, using the hypsometric equation. Differences are computed after mapping the ERA5 data to the finite-volume grids since the geopotential field is only available on the output tapes in the spectral-element runs that have been interpolated to the $\tt f09$ grid, inline.}
\label{fig:dThyps}
\end{figure}

\begin{figure}[t]
\begin{center}
         \includegraphics[width=100mm]{figs/temp_contours_diffCERES_SWCF.pdf}
\end{center}
\caption{1979-1998 Northern Hemisphere summer shortwave cloud forcing bias, relative to the CERES-EBAF gridded dataset. Shortwave cloud forcing is defined as the difference between all-sky and clear-sky net shortwave fluxes at the top of the atmosphere. Differences are computed after mapping all model output to the $1^{\circ}$ CERES-EBAF grid.}
\label{fig:SWCF}
\end{figure}

@@

\begin{figure}[t]
\begin{center}
         \includegraphics[width=130mm]{figs/temp_contours_diffCALIPSOdiffCERESdiffRACMO.pdf}
\end{center}
\caption{1979-1998 northern hemisphere summer (top) total cloud fraction bias relative to the CALIPSO dataset and incident shortwave radiation bias (W/m2), computed as the difference (middle) from CERES, and (bottom) RACMO2.3p2 dataset. The CALIPSO and CERES differences are found by mapping the model output to the $1^{\circ}$ grid, and differences in the bottom panel are computed after mapping the RACMO2.3p2 dataset to the individual model grids. Note that the averaging period for the CALIPSO-GOCCP and CERES-EBAF panels, 2006-2017 and 2003-2020, respectively, are different from the averaging period for the model results.}
\label{fig:FSDS}
\end{figure}

\begin{figure}[t]
\begin{center}
         \includegraphics[width=130mm]{figs/temp_contours_diffRACMO}
\end{center}
\caption{1979-1998 (top) annual precipitation and (bottom) ice/snow melt biases relative to RACMO2.3p2, evaluated on the native model grids. The precipitation biases are expressed as fractional changes, whereas the melt biases are absolute changes (mm/yr). In the bottom panel, the \citeA{RM2012GRL} basin boundaries are shown in grey for each model grid. Note that Figure~\ref{fig:basin} uses the basin boundaries for the two common ice masks, shown in the $\tt f19$ and $\tt ne30pg2$ panels, in computing the basin-scale integrals. Blue lines in the $\tt f19$ panel show the location of the two transects plotted in Figure~\ref{fig:ztrans}..}
\label{fig:prect}
\end{figure}

@@

\begin{figure}[t]
\begin{center}
         \includegraphics[width=100mm]{figs/temp_tseries_GRIS.pdf}
\end{center}
\caption{Time-series of annual (solid+liquid) precipitation (top) and annual runoff (bottom) integrated over the Greenland Ice Sheet for all six simulations and compared to the RACMO datasets. The time-series were generated using the common ice mask approach, which results in up to 4 ensembles, with the mean value given by the solid line and shading spanning the extent of the ensemble members.}
\label{fig:tseries}
\end{figure}

@@

\begin{figure}[t]
\begin{center}
         \includegraphics[width=100mm]{figs/temp_tseries_BASIN.pdf}
\end{center}
\caption{1979--1998 basin integrated components of the SMB; (top) precipitation, (middle) ice/snow melt and (bottom) ice/snow melt estimated from the PDD method. Whiskers span the max/min of the four ensemble members generated from the common-ice-mask approach. Basin definitions are after \citeA{RM2012GRL}, and are found on the common ice masks using a nearest neighbor approach, and shown in Figure~\ref{fig:prect}.}
\label{fig:basin}
\end{figure}

@@

\begin{figure}[t]
\begin{center}
\begin{tabular}{cccc}
         \includegraphics[width=130mm]{figs/temp_zsects.pdf}
\end{tabular}
\end{center}
\caption{Model surface elevation along the (top) K-transect, and (bottom) a transect spanning the central dome down to the Kangerlussuaq glacier in southeast Greenland, for all model grids. The reference surface (GMTED) is a 1 km surface elevation dataset used for generating the CAM topographic boundary conditions. }
\label{fig:ztrans}
\end{figure}

@@

\begin{figure}[t]
\begin{center}
\begin{tabular}{cccc}
         \includegraphics[width=80mm]{figs/temp_xy_diffRACMO_melt.pdf}
\end{tabular}
\end{center}
\caption{Fractional melt bias over the GrIS, computed relative to the RACMO datasets using the common ice mask approach, and conditionally sampled by grid cell ice fraction provided by the common ice masks. Solid lines are the mean of the distribution with $\pm$ one standard deviation expressed by shading.}
\label{fig:bias}
\end{figure}

>>>>

\begin{keypoints}
\item The CESM2.2 release includes several enhancements to the spectral-element dynamical core, including two Arctic refined mesh configurations.
\item Uniform resolution grids degrade the Greenland Ice Sheet mass balance compared to equivalent resolution latitude-longitude grids.
\item The refined Arctic meshes substantially improve the surface mass balance over all low resolution grids.
\end{keypoints}

@@

\section*{Plain Language Summary}

The mass balance of Earth's big ice sheets is crucially important for understanding controls on global sea-level rise. However, the scale of the processes needed to represent ice sheet mass balance is challenging to resolve in conventional Earth System Models. This study evaluates the ability of different grids and atmospheric solvers (i.e, the dynamical core) in CESM2.2 to resolve the surface mass balance (SMB) of the Greenland Ice Sheet. We show that the ongoing transition away from latitude longitude grids, towards quasi-uniform unstructured grids in CESM2.2 leads to a degradation of the simulated SMB. Two variable-resolution grids with enhanced resolution over Greenland are developed and incorporated into the release of CESM2.2, which substantially improves the SMB over the latitude longitude grids.
%[ enter your Plain Language Summary here or delete this section]

@@

\begin{figure}[t]
\begin{center}
%\begin{tabular}{cc}
         \includegraphics[width=100mm]{grid-uniform}
         %\includegraphics[width=50mm]{grid-f19.pdf}&
         %\includegraphics[width=50mm]{grid-f09.pdf}\\
         %\includegraphics[width=50mm]{grid-ne30pg2.pdf}&
         %\includegraphics[width=50mm]{grid-ne30pg3.pdf} \\
%\end{tabular}
\end{center}
\caption{Computational grids for the uniform $1^{\circ}-2^{\circ}$ grids in this study. Grids names after Table~\ref{tbl:table1}, (a) $\tt f19$, (b) $\tt f09$, (c) $\tt ne30pg2$ and (d) $\tt ne30pg3$.}
\label{fig:uni-grids}
\end{figure}

\begin{figure}[t]
\begin{center}
%\begin{tabular}{ccc}
         \includegraphics[width=140mm]{grid-varres.pdf}
         %\includegraphics[width=40mm]{grid-ARCTIC.pdf}&
         %\includegraphics[width=40mm]{grid-ARCTICGRIS.pdf}&
         %\includegraphics[width=40mm]{grid-CONUS.pdf} \\
%\end{tabular}
\end{center}
\caption{Variable-resolution grids available in CESM2.2, (a) $\tt Arctic$, (b) $\tt Arctic-GrIS$ and (c) $\tt CONUS$. Note what is shown is the element grid; the computational grid has $3\times3$ independent grid points per element.}
\label{fig:vr-grids}
\end{figure}

@@

\begin{figure}[t]
\begin{center}
         \includegraphics[width=100mm]{temp_grisres.pdf}
\end{center}
\caption{The spatial properties of the GrIS as represented by different grids in this study. (Left) GrIS area error, computed as the relative differences from a 4-km dataset used to create the CESM ice masks, (right) approximate average grid spacing over GrIS.}
\label{fig:grisdx}
\end{figure}

@@

\begin{figure}[t]
\begin{center}
         \includegraphics[width=130mm]{temp_dhgt_panel_STEND_CLUBB-lores.pdf}
\end{center}
\caption{1979-1998 annual mean temperature (left column) and CLUBB temperature tendencies (right column) in zonal mean height space, expressed as differences between the various $1^{\circ}-2^{\circ}$ grids. The thick grey and magenta lines are the tropopause for the control run and the test run, respectively.}
\label{fig:dT-lores}
\end{figure}

@@

\begin{figure}[t]
\begin{center}
         \includegraphics[width=130mm]{temp_dhgt_panel_STEND_CLUBB-hires.pdf}
\end{center}
\caption{As in Figure~\ref{fig:dT-lores} but for the short-time-step experiment and the VR grids. The fields plotted are for the climatological northern hemisphere summer. We focus on summer because that is when the resolution response is largest, and the refined regions are located in the northern hemisphere.}
\label{fig:dT-hires}
\end{figure}

@@

\begin{figure}[t]
\begin{center}
         \includegraphics[width=100mm]{temp_contours_diffERA5_Thyps.pdf}
\end{center}
\caption{1979-1998 lower troposphere, northern hemisphere summer virtual temperature biases, computed as the difference from ERA5. Lower troposphere layer mean virtual temperature is derived from the 1000 hPa - 500h Pa geopotential thickness, using the hypsometric equation. Differences are computed after mapping the ERA5 data to the finite-volume grids since the geopotential field is only available on the output tapes in the spectral-element runs that have been interpolated to the $\tt f09$ grid, inline.}
\label{fig:dThyps}
\end{figure}

\begin{figure}[t]
\begin{center}
         \includegraphics[width=100mm]{temp_contours_diffCERES_SWCF.pdf}
\end{center}
\caption{1979-1998 Northern Hemisphere summer shortwave cloud forcing bias, relative to the CERES-EBAF gridded dataset. Shortwave cloud forcing is defined as the difference between all-sky and clear-sky net shortwave fluxes at the top of the atmosphere. Differences are computed after mapping all model output to the $1^{\circ}$ CERES-EBAF grid.}
\label{fig:SWCF}
\end{figure}

@@

\begin{figure}[t]
\begin{center}
         \includegraphics[width=130mm]{temp_contours_diffCALIPSOdiffCERESdiffRACMO.pdf}
\end{center}
\caption{1979-1998 northern hemisphere summer surface incident shortwave radiation bias (W/m2), computed as the difference (top) from CERES, and (middle) RACMO2.3p2 dataset, and the (bottom) total cloud fraction bias relative to the CALIPSO dataset. CALIPSO and CERES differences are found by mapping the model output to the $1^{\circ}$ grid, and differences in the bottom panel are computed after mapping the RACMO2.3p2 dataset to the individual model grids. Note that the averaging period for the CALIPSO-GOCCP and CERES-EBAF panels, 2006-2017 and 2003-2020, respectively, are different from the averaging period for the model results.}
\label{fig:FSDS}
\end{figure}

\begin{figure}[t]
\begin{center}
         \includegraphics[width=130mm]{temp_contours_diffRACMO}
\end{center}
\caption{1979-1998 (top) annual precipitation and (bottom) ice/snow melt biases relative to RACMO2.3p2, evaluated on the native model grids. The precipitation biases are expressed as fractional changes, whereas the melt biases are absolute changes (mm/yr). In the bottom panel, the \citeA{RM2012GRL} basin boundaries are shown in grey for each model grid. Note that Figure~\ref{fig:basin} uses the basin boundaries for the two common ice masks, shown in the $\tt f19$ and $\tt ne30pg2$ panels, in computing the basin-scale integrals. Blue lines in the $\tt f19$ panel show the location of the two transects plotted in Figure~\ref{fig:ztrans}..}
\label{fig:prect}
\end{figure}

@@

\begin{figure}[t]
\begin{center}
         \includegraphics[width=100mm]{temp_tseries_GRIS.pdf}
\end{center}
\caption{Time-series of annual (solid+liquid) precipitation (top) and annual runoff (bottom) integrated over the Greenland Ice Sheet for all six simulations and compared to the RACMO datasets. The time-series were generated using the common ice mask approach, which results in up to 4 ensembles, with the mean value given by the solid line and shading spanning the extent of the ensemble members.}
\label{fig:tseries}
\end{figure}

@@

\begin{figure}[t]
\begin{center}
         \includegraphics[width=100mm]{temp_tseries_BASIN.pdf}
\end{center}
\caption{1979--1998 basin integrated components of the SMB; (top) precipitation, (middle) ice/snow melt and (bottom) ice/snow melt estimated from the PDD method. Whiskers span the max/min of the four ensemble members generated from the common-ice-mask approach. Basin definitions are after \citeA{RM2012GRL}, and are found on the common ice masks using a nearest neighbor approach, and shown in Figure~\ref{fig:prect}.}
\label{fig:basin}
\end{figure}

@@

\begin{figure}[t]
\begin{center}
\begin{tabular}{cccc}
         \includegraphics[width=130mm]{temp_zsects.pdf}
\end{tabular}
\end{center}
\caption{Model surface elevation along the (top) K-transect, and (bottom) a transect spanning the central dome down to the Kangerlussuaq glacier in southeast Greenland, for all model grids. The GMTED reference surface is a 1 km surface elevation dataset \cite{GMTED2010} used for generating the CAM topographic boundary conditions. }
\label{fig:ztrans}
\end{figure}

@@

\begin{figure}[t]
\begin{center}
\begin{tabular}{cccc}
         \includegraphics[width=80mm]{temp_xy_diffRACMO_melt.pdf}
\end{tabular}
\end{center}
\caption{Fractional melt bias over the GrIS, computed relative to the RACMO datasets using the common ice mask approach, and conditionally sampled by grid cell ice fraction provided by the common ice masks. Solid lines are the mean of the distribution with $\pm$ one standard deviation expressed by shading.}
\label{fig:bias}
\end{figure}
